Analysis Script

### 2025-04-04 code/analysis-code/statisitical-analysis.qmd
Fixes

```{r}
# Some quick data fixing before analysis as some models cannot handle negatives (contextually ok, more interested in NEW cases)
data_clean <- data
data_clean$new_cases[data_clean$new_cases < 0] <- 0
data_clean$new_cases[is.na(data_clean$new_cases)] <- 0
data_clean$incidence_rate[is.na(data_clean$incidence_rate)] <- 0

data_clean <- na.omit(data_clean)

# In my intial analyses, there were an absurd amount of zeroes; this was most likely due to the data sets capturing days prior to the pandemic.
last_case_date <- as.Date("2022-01-01")
first_case_date <- as.Date("2020-03-02")

df_clean <- data_clean %>%
  filter(date >= first_case_date) %>%
  filter(date <= last_case_date) %>%
  group_by(county)

df_clean <- df_clean %>%
  as_tibble() %>%
  select(-geometry)

df_clean <- df_clean %>% 
  mutate(new_cases = as.integer(df_clean$new_cases))

class(df_clean$new_cases)
```

@@

lm_fit <- lm_wf %>%
  add_model(lm_spec) %>%
  fit(data = df_clean)

@@

lm_preds<- predict(lm_fit, new_data = df_clean)

lm_metrics <- tibble(truth = df_clean$new_cases, predicted = lm_preds$.pred) %>%
  metrics(truth = truth, estimate = predicted)
lm_metrics
```

@@

lasso_fit <- lasso_wf %>%
  add_model(lasso_spec) %>%
  fit(data = df_clean)

@@

lasso_preds<- predict(lasso_fit, new_data = df_clean)

lasso_metrics <- tibble(truth = df_clean$new_cases, predicted = lasso_preds$.pred) %>%
  metrics(truth = truth, estimate = predicted)
lasso_metrics
```

@@

# 5 fold crossvalidation repeated 5 times
# Note: Limitation of computing power so doing rolling window would not work
cv_folds <- vfold_cv(data, v = 5, repeats = 5)
plan(multisession, workers = 4)
# Tune the forest
forest_tune_res <- tune_grid(
  forest_wf %>% add_model(forest_spec),
  resamples = cv_folds,  
  grid = forest_grid,
  metrics = metric_set(rmse, rsq),
  control = control_grid(parallel_over = "everything")
)

@@

best_params <- select_best(forest_tune_res, metric = "rmse")  # Or "rsq"
best_params

@@

# Calculate performance metrics (RMSE, R², etc.)
forest_rmse <- forest_preds_final %>%
  metrics(truth = truth, estimate = predicted)

>>>>

@@

lm_fit <- lm_wf %>%
  add_model(lm_spec) %>%
  fit(data = data)

@@

lm_preds<- predict(lm_fit, new_data = data)

lm_metrics <- tibble(truth = data$new_cases, predicted = lm_preds$.pred) %>%
  metrics(truth = truth, estimate = predicted)
lm_metrics
```

@@

lasso_fit <- lasso_wf %>%
  add_model(lasso_spec) %>%
  fit(data = data)

@@

lasso_preds<- predict(lasso_fit, new_data = data)

lasso_metrics <- tibble(truth = data$new_cases, predicted = lasso_preds$.pred) %>%
  metrics(truth = truth, estimate = predicted)
lasso_metrics
```

@@

# 5 fold crossvalidation repeated 5 times
# Note: Limitation of computing power so doing rolling window would not work
cv_folds <- vfold_cv(data, v = 5, repeats = 5)
plan(multisession, workers = 4)

# Tune the forest
# This step took 20-ish minutes
forest_tune_res <- tune_grid(
  forest_wf %>% add_model(forest_spec),
  resamples = cv_folds,  
  grid = forest_grid,
  metrics = metric_set(rmse, rsq),
  control = control_grid(parallel_over = "everything")
)

@@

best_params <- select_best(forest_tune_res, metric = "rmse")
best_params

@@

# Calculate performance metrics
forest_rmse <- forest_preds_final %>%
  metrics(truth = truth, estimate = predicted)
